---
title: "Bloomers work flow example"
author: "Ona Deulofeu Capo"
date: "2023-05-11"
output: github_document

---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Import functions and upload libraries
```{r}
library(devtools)
#install_github(repo = "EcologyR/Bloomers")
library(tidyverse)
library(vegan)
library(Bloomers)

source('../../R/community_evenness.R')
source('../../R/get_anomalies.R') #la poso així perquè l'he actualizat i com que no he compilat el paquet potser no funciona actualitzada 
source('../../R/calculate_relative_abundance.R')
source('../../R/blooming_event_summary.R')
source('../../R/find_asv_with_anomalies.R')
source('../../R/compute_bray_curtis_dissimilariy.R')
```

# Get example data
```{r}
#load("../../data/bloomersdata.rda")
setwd('../../')
data <- prepare_example_data()
str(data)
#head(bloomersdata)
```

# Preprocessing data 
## Calculate relative abundances
```{r}
asv_tab_l_rel_abund <- data$asv_tab_l |>
  calculate_rel_abund(group_cols = sample_id)
head(asv_tab_l_rel_abund)
```
## Calculate pseudoabundandances
Pseudoabundances are relative abundances multiplied by total abundance data for each sample.
```{r}
asv_tab_pseudoabund <- asv_tab_l_rel_abund %>%
  calculate_pseudoabund(abund_data = data$abund_data,  by_ =  c('sample_id', 'sampling_site'),
                        rel_abund = as.numeric(relative_abundance),
                        total_abund = as.numeric(mean_total_bac))
head(asv_tab_pseudoabund)
```
# Calculate diversity parammeters
We calculate different diversity parameters to check if the blooming events detected have an effect on the community structure
## Community eveness
```{r}
community_eveness <- asv_tab_pseudoabund %>%
  select(sample_id, reads, asv_num) %>%
  as_tibble() %>%
  group_by(sample_id) %>%
  dplyr::summarize(community_eveness_result = community_evenness(abundances = reads, index = "Pielou"))
head(community_eveness)
```
## Community Bray-Curtis dissimilarity
```{r}
bray_curtis <- dissimilarity_matrix(data = asv_tab_l_rel_abund, sample_id_col = sample_id, values_cols_prefix = 'HFW')
head(bray_curtis)
```
# Discover anomalies
## For each ASVs based on relative abundances and pseudoabundances
```{r}
z <- asv_tab_pseudoabund %>%
  as_tibble() |>
  group_by(asv_num) |>
  dplyr::summarize(anomalies_ab = get_anomalies(values = pseudoabundance, plotting = FALSE)[c(1,2)],
                   anomalies_ra = get_anomalies(values = relative_abundance, plotting = FALSE)[c(1,2)])
z
```
## For community eveness and bray curtis dissimilarity
```{r}
# z_diversity <- bray_curtis |>
#   dplyr::right_join(community_eveness, by = join_by('sample_id')) |> 
#   #ungroup() %>%
#   #group_by(sample_id) %>%
#   dplyr::reframe(anomalies_bray = get_anomalies(values = bray_curtis_result, plotting = FALSE)[c(1,2,3)],
#                    anomalies_eveness = get_anomalies(values = community_eveness_result, plotting = FALSE)[c(1,2,3)])
# 
# 
# z_diversity %>%
#   str()
```
# Summarize blooming events

```{r}
# z$asv_num
# #blooming_event_summary()
# asv_tab_pseudoabund |>
#   dplyr::filter(asv_num == 'asv1') |>
#   dplyr::slect()
# blooming_summary(, z_vector = )
```
# Filter the ASV_tab by only those ASVs that have an anomaly at some point of the dataset
```{r filter for ASVs that have anomalies}
asv_anom <- find_asv_with_anomalies(anomalies_result = z, anomaly_in1 = anomalies_ab, anomaly_in2 = anomalies_ra, logic1 = TRUE, logic2 = TRUE, asv_col = asv_num)

asv_tab_pseudoabund_filt <- asv_tab_pseudoabund |>
  pivot_longer(cols = c(pseudoabundance, relative_abundance), values_to = 'abundance_value', names_to = 'abundance_type') |>
  filter(asv_num %in% asv_anom)

```

## Now, we plot those ASVs that have anomalies detected
```{r plotting anomal ASVs}
asv_tab_pseudoabund_filt |> 
  ggplot(aes(sample_id, abundance_value))+
  geom_point()+
  facet_grid(abundance_type ~ asv_num, scales = 'free')+
  theme_bw()

```

